---
title: "CUNY SPS DATA 621 - CTG5 - HW4"
author: "Betsy Rosalen, Gabrielle Bartomeo, Jeremy O'Brien, Lidiia Tronina, Rose Koh"
date: "April 24th, 2019"
output:
    bookdown::pdf_document2:
        toc: true
        toc_depth: 2
        number_sections: true
        fig_width: 8
        fig_height: 6
        fig_caption: true
        includes:  
            in_header: ./source/figure_placement.tex
        highlight: haddock
        df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy = FALSE, echo=FALSE, message=FALSE, warning=FALSE)
chooseCRANmirror(graphics=FALSE, ind=1)
source("./source/libs.R")
source("./source/script.R")
source("./source/captioner.R")
#set.seed(123)
```

\newpage

# DATA EXPLORATION

In this assignment, we explore, analyze and model a dataset containing 8,161 rows and 25 columns.  Of all 25 features, 14 are discrete and 11 are continuous.  There are total 970 missing values out of 204,025 observations.

We will build a logistic and multiple linear regression that will determine the followings:
•	Predict the probability that a person will crash their car
•	Amount of money it will cost if the person does crash their car

We will be able to develop insurance rates based on a number of predictors such as income, age, distance to work, and how long they have been customers, etc.

In the training dataset, there are 23 predictors and 2 response variables – one is binary value that indicates whether claim was made and the other is numerical value indicating the cost of claim.

```{r t1}
knitr::kable(variable_descriptions, caption="Data Dictionary")
```

The response variable shows appropriate distribution in the training data.  We confirm that for the number of target flags are 0 equals the target amount 0.


```{r fig.cap = "Predictor(numeric) Histogram"}
plot_histogram(train.num)
```
```{r fig.cap="Predictor(num) Distributions by TARGET_FLAG"}
hist.new
```

```{r}
outlier.boxplot
```

```{r}
boxplots
```

```{r fig.cap = "Predictor(categorical) distribution plot"}
plot_bar(train.disc)
```

```{r fig.cap="Predictor variables scatter plat with TARGET_AMOUNT"}
plot_scatterplot(train.num.a, "TARGET_AMT")
```

```{r fig.cap="missing data plot"}
plot_missing(train)
```

\newpage


# DATA PREPARATION

\newpage


# BUILD MODELS

\newpage


# SELECT MODELS

\newpage


# Appendix

The appendix is available as script.R file in `project4_insurance` folder.

https://github.com/betsyrosalen/DATA_621_Business_Analyt_and_Data_Mining

```
```